---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

### Load the CSV File and view first serveral rows 
```{r echo=TRUE}
activity <- read.csv("~/PA1_QIAN/activity.csv")
```

### Format date
```{r echo=TRUE}
activity$date<-as.POSIXct(activity$date,format="%Y-%m-%d") 
head(activity)
```

## What is mean total number of steps taken per day?

### Total steps taken per day
```{r echo=TRUE}
#Daily total number of steps by removing missing values
total_daily<-aggregate(activity$steps, by=list(activity$date), FUN=sum,na.rm=TRUE) 
#Rename columns in table
colnames(total_daily)<-c('date','total_steps')  
total_daily
```

### Histogram of daily total steps taken
```{r echo=TRUE}
library(ggplot2)
qplot(total_daily$total_steps, geom="histogram",
      fill=I("blue"), 
      col=I("black"), 
      main = "Histogram of daily total steps taken", 
      xlab = "Daily total steps taken") 
```


### Mean and median of steps taken per day
```{r echo=TRUE}
library(doBy)
summaryBy(steps ~ date, data = activity, 
          FUN = list(mean,median))
```


## What is the average daily activity pattern?

### Average number of steps taken of the 5-minute interval across all days
```{r echo=TRUE}
step_mean_Interval<-aggregate(formula = steps~interval, 
                              data = activity,
                              FUN = mean)
colnames(step_mean_Interval)<-c('interval','Average_steps')
```

### Time series plot 
```{r echo=TRUE}
library("lattice")
xyplot(Average_steps~interval,
       data = step_mean_Interval,
       type = "l")
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo=TRUE}
# The 5-minute interval contains the maximum number of steps is 835
subset(step_mean_Interval,step_mean_Interval$Average_steps==max(step_mean_Interval$Average_steps))
```


## Imputing missing values

### Missing values

#### Missing values for steps
```{r echo=TRUE}
table(is.na(activity$steps))
```

#### Missing values for date
```{r echo=TRUE}
table(is.na(activity$date))
```

#### Missing values for interval
```{r echo=TRUE}
table(is.na(activity$interval))
```

### Replace missing values
#### Merge raw activity data with steps interval average data
```{r echo=TRUE}
steps_imputing<-merge(activity,step_mean_Interval,by="interval")
#Rename columns in table
colnames(steps_imputing)<-c('interval','steps_RAW','date','steps_mean')
head(steps_imputing)
```
#### Replace missing steps with average steps and view first several rows
```{r echo=TRUE}
#Initialise the imputing column with original data
steps_imputing$steps_imputing<-steps_imputing$steps_RAW

#Replace missing steps with average steps
steps_imputing$steps_imputing[is.na(steps_imputing$steps_RAW)==TRUE]<-steps_imputing$steps_mean[is.na(steps_imputing$steps_RAW)==TRUE]

head(steps_imputing)
```

### Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r echo=TRUE}
activity_no_missing_data<-steps_imputing
```

### The mean and median total number of steps taken per day
```{r echo=TRUE}
library(doBy)
activity_no_missing_data_Daily<-summaryBy(steps_imputing ~ date, data = activity_no_missing_data, 
                                          FUN = list(sum,mean,median))
activity_no_missing_data_Daily
```

### Histogram of the total number of steps taken each day before and after Imputing missing values
```{r echo=TRUE}
#Histogram of daily total steps taken after Imputing missing values
library(ggplot2)
qplot(activity_no_missing_data_Daily$steps_imputing.sum, geom="histogram",
      fill=I("blue"), 
      col=I("black"), 
      main = "Histogram of daily total steps taken after Imputing missing values", 
      xlab = "Daily total steps taken") 

#Histogram of daily total steps taken before Imputing missing values
library(ggplot2)
qplot(total_daily$total_steps, geom="histogram",
      fill=I("green"), 
      col=I("black"), 
      main = "Histogram of daily total steps taken before Imputing missing values", 
      xlab = "Daily total steps taken") 


```

## Are there differences in activity patterns between weekdays and weekends?

### Add weekday and weekday/weekedn flags to the data
```{r echo=TRUE}
activity_no_missing_data$date<-as.POSIXct(activity_no_missing_data$date,format="%Y-%m-%d")
#Add weekday flags and view first several rows
activity_no_missing_data$weekday<-weekdays(activity_no_missing_data$date)
head(activity_no_missing_data)

#Add weekday/weekend flags and view first several rows
activity_no_missing_data$weekday_weekend_Flag<-"weekday"
activity_no_missing_data$weekday_weekend_Flag[activity_no_missing_data$weekday %in% c("Saturday", "Sunday")]<-"weekend"
head(activity_no_missing_data)
```

### 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days
```{r echo=TRUE}
step_mean_weekday_weekend<-aggregate(formula = steps_imputing~interval+weekday_weekend_Flag, 
                                     data = activity_no_missing_data,
                                     FUN = mean)
colnames(step_mean_weekday_weekend)<-c('interval','weekday_weekend_Flag','Average_steps')
```

### Panel plot containing a time series plot for weekday and weekend
```{r echo=TRUE}
library("lattice")
xyplot(Average_steps~interval|weekday_weekend_Flag,
       data = step_mean_weekday_weekend,
       type = "l")
```